# Logistic Regression 

*In this chapter, we learnt how to read data from the internet and combine two data sets which do not have unique identifiers.
We also learnt how to explore data with bar and box plots. Lastly, we learnt how to model our data using a linear regression model.*

<!-- ##reading the data -->
#1. Reading data
```{r read data_1, eval=TRUE, echo=FALSE, warning=FALSE}
alc = read.table(url("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/alc.txt "), sep=",",header=TRUE)
dim(alc)
str(alc)
```
### Summary of data
This dataset has 382 rows and 35 variables. 
There is general information about the students: school they belong to, sex, age, address and family size. There is also information about their parents,
allocation of hours towards travel to school and studytime, some personal information about what the student is otherwise occupied with, health status (including alcohol consumption) and grades for Math and Portugese across 2 periods.

#2. Hypothesis
Hypotheses to be tested
1. alcohol use is associated with the quality of family relationships.
2. alcohol use is associated with student's freetime.
3. alcohol use is associated with health.
4. alcohol use is associated with sex.

#3. Variable means
```{r tabulate data, eval=TRUE, echo=FALSE, warning=FALSE}

# access the tidyverse libraries tidyr, dplyr, ggplot2
library(tidyr) 
library(dplyr)
library(ggplot2)

# glimpse at the alc data
#glimpse(alc)

alc %>% group_by(high_use) %>% summarise(count = n(),mean_famrel=(mean(famrel)),mean_freetime=(mean(freetime)), mean_health=(mean(health)) )


```
### Summary
Here, we see that there are 270 cases of low alcohol use and 112 cases of high alcohol use (more than 2 on a scale of 1-5).
freetime - free time after school (numeric: from 1 - very low to 5 - very high) 
famrel - quality of family relationships (numeric: from 1 - very bad to 5 - excellent) 
health - current health status (numeric: from 1 - very bad to 5 - very good) 

In the group of high alcohol use, the mean of quality of family relationships is 3.7 
and for the group with low alcohol use, the mean of quality of family relationships 
is 4.0. In general, it can be surmised that low alcohol use may be associated with better 
family relationships.

Amongst students with low alcohol use, there is less free time (mean = 3.12) compared to 
students with higher alcohol use who seem to have more free time (mean = 3.46). Interestingly,
low alcohol consumption students seem to enjoy worse health (mean = 3.5) than students who drink more
(mean = 3.72)


#4. Variable means by sex
```{r tabulate data2, eval=TRUE, echo=FALSE, warning=FALSE}
alc %>% group_by(sex,high_use) %>% summarise(count = n(),mean_famrel=(mean(famrel)),mean_freetime=(mean(freetime)), mean_health=(mean(health)) )
```
### Summary
With regards to sex, 157 females consume low amounts of alcohol and 41 females consume high amounts
of alcohol. In the group of females who drink more, family relationship is worse ( 3.73), they have more freetime 
(mean=3.34) and they enjoy better health (3.39) as compared to females who drink less.

In the male group, there are 113 males who drink less and 71 males who drink more. The group that drinks more
has less quality family relationships (mean=3.77), more freetime (mean=3.52) and also enjoy better healht 
(3.91) as compared to males  who drink less. Males with high alcohol use have more free time and better family relationships 
and enjoy better health than females with high alcohol use.

#5. Boxplots for selected variables
```{r visualize data_1, eval=TRUE, echo=FALSE, warning=FALSE}
##################################################box plot#############################################################################

par(mfrow = c(1, 3), mar = c(3, 1, 1, 1))
#1)alcohol_use and failures
# initialize a plot of alcohol_use and famrel
g1 <- ggplot(alc, aes(x = high_use, y = famrel, col=sex))
# define the plot as a boxplot and draw it
g1 + geom_boxplot() + ylab("family relationship")  + ggtitle("Family relationship by alcohol consumption and sex") + stat_summary(fun.y = "mean", geom = "text", label="----", size= 10, color= "black")

#2)alcohol_use and parents co-habitation status
# initialise a plot of alcohol_use and students freetime
g2 <- ggplot(alc, aes(x = high_use, y = freetime, col=sex)) + ggtitle("Student freetime by alcohol consumption and sex")
# define the plot as a boxplot and draw it
g2 + geom_boxplot() + ylab("Student freetime") + stat_summary(fun.y = "mean", geom = "text", label="----", size= 10, color= "black")

#3)alcohol_use and health
# initialise a plot of alcohol_use and health
g2 <- ggplot(alc, aes(x = high_use, y = health, col=sex)) + ggtitle("Student health by alcohol consumption and sex")
# define the plot as a boxplot and draw it
g2 + geom_boxplot() + ylab("health") + stat_summary(fun.y = "mean", geom = "text", label="----", size= 10, color= "black")

################################################bar plot###############################################################################


```




### Summary
The median for quality of family relationships is the same for both males and females 
across both groups of alcohol use (median = 4). The upper quartile value is 5 in both sexes
for students with low alcohol use and 4 for students with high alcohol use.
The lowest value for family relationship (ignoring the outliers) is 3 for students with low alcohol
consumption and 2 for students with high alcohol consumption. 

The median for student freetime is 3 for males and females for students with low alcohol use and 3 
for females and 4 for males in the group with higher alcohol use. 
The upper quartile value is 5 in both sexes.


The median for student health is 4 for students with high and low alcohol use.
Mean health is higher for students with high alcohol comsumption.

#6. Barplots for selected variables

```{r barplots, eval=TRUE, echo=FALSE, warning=FALSE}
# use gather() to gather columns into key-value pairs and then glimpse() at the resulting data
# initialize a plot of 'high_use'
g2 <- ggplot(data = alc,aes(x=famrel))
# draw a bar plot of high_use by sex
g2+geom_bar()+facet_wrap("sex")

g2 <- ggplot(data = alc,aes(x=freetime))
# draw a bar plot of high_use by sex
g2+geom_bar()+facet_wrap("sex")

g2 <- ggplot(data = alc,aes(x=health))
# draw a bar plot of high_use by sex
g2+geom_bar()+facet_wrap("sex")

g2 <- ggplot(data = alc,aes(x=high_use))
# draw a bar plot of high_use by sex
g2+geom_bar()+facet_wrap("sex")

```


### Summary
The pattern for family relationship across sexes is the same.
For freetime, the smallest number of females have the most freetime (5) and the smallest number of males have the lowest freetime (1).
For health, the highest number of individuals have the healthiest status in both sexes. Across both sexes, there are more individuals with low alcohol use.

#7. Regression model with selected variables (famrel + freetime + health + G1)
```{r model data1, eval=TRUE, echo=FALSE, warning=FALSE}

# find the model with glm()
m <- glm(high_use ~ famrel + freetime + health + G1, data = alc, family = "binomial")

# print out a summary of the model
summary(m)

# print out the coefficients of the model
coef(m)

# compute odds ratios (OR)
OR <- coef(m) %>% exp

# compute confidence intervals (CI)
CI=exp(confint(m))

# print out the odds ratios with their confidence intervals
cbind(OR, CI)
```
### Summary
Family relationship and first period scores are negatively associated with alcohol consumption and student freetime is positively associated with alcohol consumption.
There is no association between student health and alcohol consumption.


#8. Regression model with selected variables (famrel + freetime + sex + G1)
```{r model data2, eval=TRUE, echo=FALSE, warning=FALSE}

# find the model with glm()
m <- glm(high_use ~ famrel + freetime + sex + G1, data = alc, family = "binomial")

# print out a summary of the model
summary(m)

# print out the coefficients of the model
coef(m)

# compute odds ratios (OR)
OR <- coef(m) %>% exp

# compute confidence intervals (CI)
CI=exp(confint(m))

# print out the odds ratios with their confidence intervals
cbind(OR, CI)
```

### Summary
Family relationships and period 1 scores are negatively associated with alcohol use 
i.e. the lower the quality of family relationships, the higher the alcohol use. The most significant association is sex. With sex=females as
the baseline in the intercept, the coefficient for males is 0.06 + 0.91=0.97.
In summary, family relationships, student freetime, sex and period 1 scores are are significantly associated with high alcohol use.

#9. predictions versus the actual values
```{r predict data, eval=TRUE, echo=FALSE, warning=FALSE}
# predict() the probability of high_use
m <- glm(high_use ~ famrel + freetime + G1 + sex, data = alc, family = "binomial")
probabilities <- predict(m, type = "response")
# add the predicted probabilities to 'alc'
alc <- mutate(alc, probability = probabilities)

# use the probabilities to make a prediction of high_use
alc <- mutate(alc, prediction = ( (probability>0.5)))
table(alc$high_use,alc$prediction)
# tabulate the target variable versus the predictions

g <- ggplot(alc, aes(x = probability, y = high_use,col=prediction))

# define the geom as points and draw the plot
g + geom_point()
```

### Summary
The model is more accurate at predicting high alcohol use = FALSE. Out of 270 cases that were FALSE (high alcohol use) the model predicted
20 cases to be TRUE (high alcohol use) and  240 as false. Out of 112 cases that were TRUE (high alcohol use) the model predicted
94 cases to be false (high alcohol use) and  18 as true. 

#10. Percentage predictions accuracy
```{r predict data2, eval=TRUE, echo=FALSE, warning=FALSE}




# tabulate the target variable versus the predictions
table(high_use = alc$high_use, prediction = alc$prediction) %>% prop.table() %>% addmargins()
```

### Summary
Predition error is 30% (0.25 + 0.05) of the time.
