---
title: "Exercise 5"
author: "Mahes Muniandy"
date: "24 February 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dimension Reduction

This week, we learnt how to do dimension reduction. I did not do the last part-
no 11.

## Summary of data

```{r readdata, echo=FALSE, warning=FALSE, message=FALSE}

human = read.table("human2.txt", sep="\t", quote = "", stringsAsFactors = FALSE, header=TRUE)
dim(human)

str(human)

library(FactoMineR)
library(GGally)
library(ggplot2)
library(dplyr)
library(tidyr)
library(MASS)
```

There are 154 rows and 8 variables:

 - gii_FM_edu_rat: ratio of females over males with at least secondary education
 
 - gii_FM_work_rat  : ratio of females over males who work
 
 - exp_edu          : Expected years of schooling 
 
 - life_exp         : Life expectancy at birth
 
 - GNI              : Gross National Income per capita
 
 - mat_mor_rat      : Maternal mortality ratio
 
 - adol_birth_rat   : Adolescent birth rate
 
 - fem_parlimen     : percentage of females in parlimen


## Variable correlation

```{r plot data human, eval=TRUE, message = FALSE, warnings= FALSE, echo=FALSE}
# access the GGally and ggplot2 libraries
library(GGally)
library(ggplot2)

# create a more advanced plot matrix with ggpairs(learning2014[-1])
p <- ggpairs(human, mapping = aes( alpha=0.3), lower = list(combo = wrap("facethist", bins = 20)))

# draw the plot
p

```
The dataset plot shows that most of the distributions of the variables are not normal.


```{r correlation of variables, eval=TRUE, message = FALSE, warnings= FALSE, echo=FALSE}

# calculate the correlation matrix and round it
cor_matrix<-cor(human) 

# print the correlation matrix
cor_matrix %>% round(2)  %>% print() 

library(corrplot)

cor.mtest <- function(mat, conf.level = 0.95){
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- lowCI.mat <- uppCI.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  diag(lowCI.mat) <- diag(uppCI.mat) <- 1
  for(i in 1:(n-1)){
    for(j in (i+1):n){
      tmp <- cor.test(mat[,i], mat[,j], conf.level = conf.level)
      p.mat[i,j] <- p.mat[j,i] <- tmp$p.value
      lowCI.mat[i,j] <- lowCI.mat[j,i] <- tmp$conf.int[1]
      uppCI.mat[i,j] <- uppCI.mat[j,i] <- tmp$conf.int[2]
    }
  }
  return(list(p.mat, lowCI.mat, uppCI.mat))
}

res1 <- cor.mtest(human,0.95)
# visualize the correlation matrix
corrplot.mixed(cor_matrix,tl.cex=0.5,p.mat = res1[[1]], insig = "blank"
,mar=c(1,0,2,0),number.cex=0.7, title="Correlation between variables")

```


The table above shows the correlations between the variables. The correlation plot below it shows the value of the above correlations with circles: the bigger the circle, the higher the correlation. Only significant (p<0.05) correlations are shown.

There is a high, significant correlation between life expectancy and 
expected education (0.79) and expected education and GNI (0.62). There 
is also a high negative correlation between expected education (-0.73)
and adolescence birth rate (-0.7). Life expectancy also has a high positive
correlation (0.63) with GNI. There is also a high negative correlation 
between life expectancy and maternal mortality rate (-0.85) and life expectancy and adolescence birth rate (-0.72). There is also a significant correlation between GNI and adolescence birth rate (-0.55) and maternal mortality rate and adolescence birth rate (0.75).
 
## PCA with unscaled data 
```{r pca unscaled, echo=FALSE, message=FALSE, warning=FALSE}
pca_human <- prcomp(human)

s<-summary(pca_human)
s
pca_pr<-round(100*s$importance[2,], digits=1)
pca_pr
pc_lab<-paste0(names(pca_pr),"(",pca_pr,"%)")
biplot(pca_human, , choices =1:2, cex=c(0.8,1),col=c("grey40","deeppink2"),xlab=pc_lab[1],ylab=pc_lab[2]) 

```
The plot shows that 1 PC is enough to capture almost 100% of the variation in the data. The biplot shows that countries that cluster together have high correlations.
 
## PCA with scaled data  
```{r pca scaled, echo=FALSE, message=FALSE, warning=FALSE}

human_scaled=scale(human)
human_scaled=as.data.frame(human_scaled)
pca_human <- prcomp(human_scaled)

s<-summary(pca_human)
s
pca_pr<-round(100*s$importance[2,], digits=1)
pca_pr
pc_lab<-paste0(names(pca_pr),"(",pca_pr,"%)")
biplot(pca_human, choices =1:2, cex=c(0.8,1),col=c("grey40","deeppink2"),xlab=pc_lab[1],ylab=pc_lab[2]) 

```
The scaled data shows that females in parlimen are not highly correlated with maternal mortality rate
as well as adoscelence birth rate. There is also negative correlation between expected education and life 
expectancy with maternal mortality ratio aand adolescence birth rate. The countries that are
closer together are more correlated.

The plot using the scaled data seems more accurate (when we refer to the correlation plot)
because the data are all compared on the same scale. With this plot, it is clear that with PC1 and 2, about 59.6% variation in data is explained.