# Logistic Regression 

*In this chapter, we learnt how to perform logistic regression*

<!-- ##reading the data -->

```{r read data, eval=TRUE, echo=FALSE, warning=FALSE}

alc = read.table(url("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/alc.txt "), sep=",",header=TRUE)
dim(alc)
str(alc)
```
This dataset has 382 rows and 35 variables. 
There is general information about the students: school they belong to, sex, age, address and family size. There is also information about their parents,
allocation of hours towards travel to school and studytime, some personal information about what the student is otherwise occupied with, health status (including alcohol consumption) and grades for Math and Portugese across 2 periods.

Hypothesis to be tested
1) alcohol use is associated with number of failures
2) alcohol use is associated with mother's education
3) alcohol use is associated with health
4) alcohol use is associated with sex
```{r tabulate data, eval=TRUE, echo=FALSE, warning=FALSE}

# access the tidyverse libraries tidyr, dplyr, ggplot2
library(tidyr) 
library(dplyr)
library(ggplot2)

# glimpse at the alc data
#glimpse(alc)

alc %>% group_by(high_use) %>% summarise(count = n(),mean_famrel=(mean(famrel)))

alc %>% group_by(high_use) %>% summarise(count = n(),mean_freetime=(mean(freetime)))

alc %>% group_by(high_use) %>% summarise(count = n(),mean_health=(mean(health)))

alc %>% group_by(high_use) %>% summarise(count = n(),mean_sex=(mean(sex)))

```
Here, we see that in both males and females, the highest number of individuals (89females and 55 males) consume the smallest amount of alcohol (1 unit).
The highest mean for family relat



```{r visualize data, eval=TRUE, echo=FALSE, warning=FALSE}
##################################################box plot#############################################################################

par(mfrow = c(1, 3), mar = c(3, 1, 1, 1))
#1)alcohol_use and failures
# initialize a plot of alcohol_use and famrel
g1 <- ggplot(alc, aes(x = high_use, y = famrel))
# define the plot as a boxplot and draw it
g1 + geom_boxplot() + ylab("family relationship")  + ggtitle("Family relationship by alcohol consumption")

#2)alcohol_use and parents co-habitation status
# initialise a plot of alcohol_use and parents co-habitation status
g2 <- ggplot(alc, aes(x = high_use, y = freetime)) + ggtitle("freetime by alcohol consumption")
# define the plot as a boxplot and draw it
g2 + geom_boxplot() + ylab("freetime")

#3)alcohol_use and health
# initialise a plot of alcohol_use and health
g2 <- ggplot(alc, aes(x = high_use, y = health)) + ggtitle("Student health by alcohol consumption")
# define the plot as a boxplot and draw it
g2 + geom_boxplot() + ylab("health")

#3)alcohol_use and sex
# initialise a plot of alcohol_use and health
g2 <- ggplot(alc, aes(x = high_use, y = sex)) + ggtitle("Student sex by alcohol consumption")
# define the plot as a boxplot and draw it
g2 + geom_boxplot() + ylab("sex")

################################################bar plot###############################################################################
# use gather() to gather columns into key-value pairs and then glimpse() at the resulting data
# initialize a plot of 'high_use'
g2 <- ggplot(data = alc,aes(x=famrel))
# draw a bar plot of high_use by sex
g2+geom_bar()+facet_wrap("sex")

g2 <- ggplot(data = alc,aes(x=freetime))
# draw a bar plot of high_use by sex
g2+geom_bar()+facet_wrap("sex")

g2 <- ggplot(data = alc,aes(x=health))
# draw a bar plot of high_use by sex
g2+geom_bar()+facet_wrap("sex")

g2 <- ggplot(data = alc,aes(x=sex))
# draw a bar plot of high_use by sex
g2+geom_bar()+facet_wrap("sex")

g2 <- ggplot(data = alc,aes(x=high_use))
# draw a bar plot of high_use by sex
g2+geom_bar()+facet_wrap("sex")


```

```{r model data, eval=TRUE, echo=FALSE, warning=FALSE}

# find the model with glm()
m <- glm(high_use ~ famrel + freetime + G1 +sex, data = alc, family = "binomial")

# print out a summary of the model
summary(m)

# print out the coefficients of the model
coef(m)

# compute odds ratios (OR)
OR <- coef(m) %>% exp

# compute confidence intervals (CI)
CI=exp(confint(m))

# print out the odds ratios with their confidence intervals
cbind(OR, CI)
```
Interpret the results and compare them to your previously stated hypothesis
```{r predict data, eval=TRUE, echo=FALSE, warning=FALSE}
# predict() the probability of high_use
m <- glm(high_use ~ famrel + freetime + G1, data = alc, family = "binomial")
probabilities <- predict(m, type = "response")

# add the predicted probabilities to 'alc'
alc <- mutate(alc, probability = probabilities)

# use the probabilities to make a prediction of high_use
alc <- mutate(alc, prediction = ( (probability>0.5)))
table(alc$high_use,alc$prediction)
# tabulate the target variable versus the predictions

g <- ggplot(alc, aes(x = probability, y = high_use,col=prediction))

# define the geom as points and draw the plot
g + geom_point()

# tabulate the target variable versus the predictions
table(high_use = alc$high_use, prediction = alc$prediction) %>% prop.table() %>% addmargins()
```